Life in the canopy: community trait assessments reveal substantial functional diversity among fern epiphytes

### 2020-01-08 ms/manuscript.Rmd
Fix climate data processing
- Write out moorea climate data for Dryad
- Use 'rh' as column name in raw data
- Fix measurement interval (15 min not 30 min) in MS, use code for start and end times instead of hard values

# Load analysis results
loadd(list = c(
  "best_fit_div_models",
  "climate_ancova_results",
  "climate_plot",
  "climate_select",
  "combined_comm_div_plots",
  "combined_cwm_plots",
  "comm",
  "correlated_evo_test",
  "correlated_evo_test_gemmae",
  "daily_mean_climate",
  "div_el_model_summaries",
  "div_t_test_results",
  "div_metrics_all",
  "fern_traits",
  "importance_heatmap",
  "pca_plot",
  "pca_results",
  "phy",
  "phylosig_binary_traits",
  "phylosig_cont_traits",
  "pic_results",
  "species_list",
  "traits_with_tree"),
  cache = canopy_cache)
```

@@

```{r climate-numbers, results="hide"}
# Count total number of days in climate data
total_days <- daily_mean_climate %>% pull(date) %>% n_distinct()

# Check start and end dates
# - start July 7, 2013
assert_that(daily_mean_climate %>% pull(date) %>% min == "2013-07-07")
# - end July 5, 2014
assert_that(daily_mean_climate %>% pull(date) %>% max == "2014-07-05")

# Count total number of sites in climate data
total_sites <-
daily_mean_climate %>%
  select(site, habit) %>%
  unique() %>%
  count(habit)
```

Temperature and RH of epiphytic and terrestrial habitats were measured using Hobo Pro v2 dataloggers with the RS3 Solar Radiation Shield (Onset Corporation, USA).
A pair of dataloggers was installed for each plot, one mounted at \circa \SI{2}{\meter} on a tree ('epiphytic datalogger'), and one at ground level ('terrestrial datalogger'), except for the Mt. Tohiea \SI{1200}{\meter} plot, which only had an epiphytic datalogger.
Temperature and RH were logged once every \SI{30}{\minute} during two survey periods: a preliminary survey from July 18, 2012 to July 7, 2013 at six plots, and a final survey from July 7, 2013 to July 5, 2014 at 15 plots.
Two plots were not included in the environmental survey.
We used the 2013–-2014 data for the analysis, with one exception: the terrestrial datalogger at the Mt. Tohiea \SI{393}{\meter} plot malfunctioned and was missing data for a substantial part of the final survey period (March 12--July 5, 2014), so we used data for the same dates from the preliminary survey period for this datalogger instead.
Three others also failed early during the final survey period, but lacked data from the preliminary period, so were excluded from analysis.
Other days missing data during the final survey period were excluded from the dataset.
The final dataset included `r total_days` days of data for `r total_sites %>% filter(habit == "epiphytic") %>% pull(n)` dataloggers and `r total_sites %>% filter(habit == "terrestrial") %>% pull(n)` terrestrial dataloggers.
We calculated vapor pressure deficit (VPD) from relative humidity and temperature using the \package{plantecophys} R package [@Duursma2015].

@@

Morphological traits of sporophytes were measured from herbarium specimens, or obtained from the literature when plant material was unavailable.
Measurements were made directly on plants in the field for a few species with large fronds.
A single leaf or rhizome per individual was measured on one to three separately collected individuals per species, and mean values were calculated for each species.
Although trait values can vary within fern species with elevation [@Kessler2007a], our study includes a broad phylogenetic sampling, so we assume intraspecific variation is outweighed by interspecies differences.
To measure specific leaf area (SLA; \si{\square\meter\per\kilo\gram}), \circa 10--12, \SI{2}{\milli\meter} or \SI{4}{\milli\meter} \diameter punches were taken of laminae between primary veins for a single leaf per individual using biopsy punches, dried at \SI{60}{\celsius} for \SI{24}{\hour}, and weighed.
For species with \textless \SI{2}{\milli\meter} between primary veins, leaf fragments excluding primary veins were obtained by dissection, dried at \SI{60}{\celsius} for \SI{24}{\hour}, weighed, and scanned.
ImageJ [@Abramoff2004] software was used to measure area per fragment.

@@

**\fig \@ref(fig:climate)**: Selected microclimatic variables of study plots along an elevational gradient from \SIrange{200}{1200}\meter on Moorea, French Polynesia. Relative humidity (RH) and temperature were recorded every \SI{15}{\minute} with dataloggers from July 7, 2013 to July 5, 2014. Vapor pressure deficit (VPD) was calculated from relative humidity and temperature as described in the methods, and overall means calculated for daily mean, minimum, maximum, and standard deviation (SD). 'Epiphytic' dataloggers were placed at \circa 2 m on trees, and 'terrestrial' dataloggers placed at ground level. Color indicates growth habit: epiphytic dataloggers in green, terrestrial dataloggers in brown. Trendlines fitted using linear models with an interaction between growth habit and elevation (see Methods); all trendlines significant at \pval \textless \num{.05}

>>>>

# Load analysis results
loadd(list = c(
  "best_fit_div_models",
  "climate_ancova_results",
  "climate_plot",
  "climate_raw",
  "climate_select",
  "combined_comm_div_plots",
  "combined_cwm_plots",
  "comm",
  "correlated_evo_test",
  "correlated_evo_test_gemmae",
  "daily_mean_climate",
  "div_el_model_summaries",
  "div_t_test_results",
  "div_metrics_all",
  "fern_traits",
  "importance_heatmap",
  "moorea_climate",
  "pca_plot",
  "pca_results",
  "phy",
  "phylosig_binary_traits",
  "phylosig_cont_traits",
  "pic_results",
  "species_list",
  "traits_with_tree"),
  cache = canopy_cache)
```

@@

```{r climate-numbers, results="hide"}
# Calculate summary statistics (number of sites, start date, end date) for raw 
# climate data for preliminary survey period (until 2013-07-07) and final survey
# period (2013-07-07 and after) separately.

prelim_climate_survey <-
  # Subset to Moorea sites used in prelim survey (Mt. Tohiea only)
  climate_raw %>%
  filter(str_detect(site, "Aorai", negate = TRUE)) %>%
  filter(date < "2013-07-07") %>%
  filter(str_detect(site, "Tohiea")) %>%
  # Only count by site name, not distinguishing between epi vs. ter dataloggers
  mutate(site = str_remove_all(site, "_epi|_ter")) %>%
  summarize(
    n_sites = n_distinct(site),
    start = min(date) %>% 
      format('%B %e, %Y') %>%
      str_replace_all("  ", " "),
    end = max(date) %>% 
      format('%B %e, %Y') %>%
      str_replace_all("  ", " ")
  )

final_climate_survey <-
  # Subset to Moorea sites used in final survey
  climate_raw %>%
  filter(str_detect(site, "Aorai", negate = TRUE)) %>%
  filter(date >= "2013-07-07") %>%
  # Only count by site name, not distinguishing between epi vs. ter dataloggers
  mutate(site = str_remove_all(site, "_epi|_ter")) %>%
  summarize(
    n_sites = n_distinct(site),
    start = min(date) %>% 
      format('%B %e, %Y') %>%
      str_replace_all("  ", " "),
    end = max(date) %>% 
      format('%B %e, %Y') %>%
      str_replace_all("  ", " ")
  )

# Count total number of days in clean climate data
total_days <- moorea_climate %>% pull(date) %>% n_distinct()

# Count total number of sites / dataloggers in clean climate data
total_sites <-
  moorea_climate %>%
  count(site, habit) %>%
  count(habit)
```

Temperature and RH of epiphytic and terrestrial habitats were measured using Hobo Pro v2 dataloggers with the RS3 Solar Radiation Shield (Onset Corporation, USA).
A pair of dataloggers was installed for each plot, one mounted at \circa \SI{2}{\meter} on a tree ('epiphytic datalogger'), and one at ground level ('terrestrial datalogger'), except for the Mt. Tohiea \SI{1200}{\meter} plot, which only had an epiphytic datalogger.
Temperature and RH were logged once every \SI{15}{\minute} during two survey periods: a preliminary survey from `r prelim_climate_survey$start` to `r prelim_climate_survey$end` at `r prelim_climate_survey$n_sites %>% english2` plots, and a final survey from `r final_climate_survey$start` to `r final_climate_survey$end` at `r final_climate_survey$n_sites %>% english2` plots.
Two plots were not included in the environmental survey.
We used the 2013–-2014 data for the analysis, with one exception: the terrestrial datalogger at the Mt. Tohiea \SI{393}{\meter} plot malfunctioned and was missing data for a substantial part (March 12--July 5, 2014) of the final survey period, so we used data for the same dates from the preliminary survey period for this datalogger instead.
Three others also failed early during the final survey period, but lacked data from the preliminary period, so were excluded from analysis.
Other days missing data during the final survey period were excluded from the dataset.
The final dataset included `r total_days` days of data for `r total_sites %>% filter(habit == "epiphytic") %>% pull(n)` epiphytic dataloggers and `r total_sites %>% filter(habit == "terrestrial") %>% pull(n)` terrestrial dataloggers.
We calculated vapor pressure deficit (VPD) from relative humidity and temperature using the \package{plantecophys} R package [@Duursma2015].

@@

Morphological traits of sporophytes were measured from herbarium specimens, or obtained from the literature when plant material was unavailable.
Measurements were made directly on plants in the field for a few species with large fronds.
A single leaf or rhizome per individual was measured on one to three separately collected individuals per species, and mean values were calculated for each species.
Although trait values can vary within fern species with elevation [@Kessler2007a], our study includes a broad phylogenetic sampling, so we assume intraspecific variation is outweighed by interspecies differences.
To measure specific leaf area (SLA; \si{\square\meter\per\kilo\gram}), \circa 10--12, \SI{2}{\milli\meter} or \SI{4}{\milli\meter} diam. punches were taken of laminae between primary veins for a single leaf per individual using biopsy punches, dried at \SI{60}{\celsius} for \SI{24}{\hour}, and weighed.
For species with \textless \SI{2}{\milli\meter} between primary veins, leaf fragments excluding primary veins were obtained by dissection, dried at \SI{60}{\celsius} for \SI{24}{\hour}, weighed, and scanned.
ImageJ [@Abramoff2004] software was used to measure area per fragment.

@@

**\fig \@ref(fig:climate)**: Selected microclimatic variables of study plots along an elevational gradient from \SIrange{200}{1200}\meter on Moorea, French Polynesia. Relative humidity (RH) and temperature were recorded every \SI{15}{\minute} with dataloggers from `r final_climate_survey$start` to `r final_climate_survey$end`. Vapor pressure deficit (VPD) was calculated from relative humidity and temperature as described in the methods, and overall means calculated for daily mean, minimum, maximum, and standard deviation (SD). 'Epiphytic' dataloggers were placed at \circa 2 m on trees, and 'terrestrial' dataloggers placed at ground level. Color indicates growth habit: epiphytic dataloggers in green, terrestrial dataloggers in brown. Trendlines fitted using linear models with an interaction between growth habit and elevation (see Methods); all trendlines significant at \pval \textless \num{.05}
